---
title: "GROW 2023 Data"
author: "Shayna Sura"
format: html
editor: visual
---



<!-- TO DO - FIGURE OUT WHY THE SEAGRASS CANOPY HEIGHT AVERAGES FOR EACH PLOT ARE NOT PLOTTING IN THE RIGHT LOCATION FOR THEIR Y_VALUES!!! -->



## Data Information

### About the data

Data were downloaded as .csv files from Google Drive.

Oyster data [link](https://docs.google.com/spreadsheets/d/1SPkAdKGaTz9FBQBvVFP4YggArjNHhA2ZsPg7ZFa85XQ/edit?pli=1#gid=0) for accessing data on Google Drive.

Seagrass data [link](https://docs.google.com/spreadsheets/d/1B67Mj2S6X3iJpDYtuD5ynjN1QRwLmFrBar1ooeU03_0/edit?pli=1#gid=811567800) for accessing data on Google Drive.


### Set up

```{r}
#| echo: false
#| warning: false
#| message: false


library(readr)
library(tidyr)
library(dplyr)
library(here)
library(ggplot2)

```


### Read in data
```{r}
#| echo: false
#| warning: false
#| message: false


seagrass_data <- read_csv(here("data/P2 Initial Seagrass - seagrass_initial_data.csv"))
oyster_size <- read_csv(here("data/P2 Initial Oysters - size.csv"))
oyster_abundance <- read_csv(here("data/P2 Initial Oysters - abundance.csv"))

# otherwise can use the ".." to move out of the R directory and then into the data directory.

```

```{r exploring our seagrass data}
#| echo: false
#| eval: false


glimpse(seagrass_data)

# the Percent_Cover and Canopy_Height_cm variables are reading in as characters instead of numeric. Why?

# there are some Percent_Cover values that are given as ranges (e.g., 0-1 or 0-5), and dashes are characters, not numbers.

# there are N/As for Canopy_Height_cm since those measurements were not taken at all time points.

```


```{r exploring our oyster data}
#| echo: false
#| eval: false


glimpse(oyster_size)
glimpse(oyster_abundance)

levels(as.factor(oyster_size$Plot_Type))
levels(as.factor(oyster_size$Plot_Label))

# we have a row for the "extra" GROW tile we deployed, but we don't want to include that in our summaries, so going to remove those rows from the data.

# we also have a plot label of "ST5" which seems like it might be a type.
## Shayna commented on this on the google sheet for Eme to double check. In the meantime, Shayna is manually changing this to ST3 in her downloaded .csv file of these data because it seems highly unlikely that it's anything else.

```



Decisions made about the data:
 - For oyster data, the data recorded for the "extra" GROW tile is removed and not included in further analyses.
 - For seagrass percent cover values that were recorded as ranges on the datasheets (e.g., 0-1 or 0-5), the values are being overwritten with the lower end of the range given so as to provide a conservative estimate of seagrass cover. So for a range of '0-5', the actual data value will be changed to '0'.

```{r}
#| echo: false



## Oyster Data
# need to remove the rows that have data on the "Extra" GROW tile.
# need to remove the rows that have NA values for the oyster abundance

oyster_size_clean <- oyster_size %>% 
  filter(Plot_Type != "Extra")

oyster_abundance_clean <- oyster_abundance %>% 
  filter(Plot_Type != "Extra") %>% 
  filter(Plot_Label != "#N/A")



## Seagrass Data
# need to convert the Percent_Cover and Canopy_Height_cm variables to numeric instead of characters.

# can also remove unnecessary columns, specifically 'entered' and 'QA/QC'

# convert the 'Date' column to date format instead of character format?

seagrass_clean <- seagrass_data %>% 
  mutate(Percent_Cover = as.numeric(Percent_Cover)) %>% 
  mutate(Canopy_Height_cm = as.numeric(Canopy_Height_cm)) %>% 
  select(-entered) %>% 
  select(-'QA/QC')

## determine which rows in Percent_Cover were converted to NAs because the data was recorded as a range and reassign those NA values to the lower value of the range reported.

ranges <- which(is.na(seagrass_clean$Percent_Cover)) # will tell us which row(s) have NAs


# seagrass_data[ranges,]


# change the NA in the seagrass_clean for these rows to the lower value of the range given in seagrass_data
for (k in ranges){
  split_values <- strsplit(seagrass_data$Percent_Cover[k], "-")
  lower_value <- sapply(split_values, function(x) as.numeric(x[1]))
  seagrass_clean$Percent_Cover[k] = lower_value
}


## double check the for-loop worked correctly...
# seagrass_clean[ranges,]


```



## Summary Statistics

### Calculating Summary Statistics

#### Seagrass Data - Percent Cover & Canopy Height
```{r}
#| echo: false


summary_calcs <- list(
  Mean = function(x) mean(x, na.rm = TRUE),
  n = function(x) sum(!is.na(x)),
  SD = function(x) ifelse(all(is.na(x)), NA, sd(x, na.rm = TRUE)),
  SE = function(x) ifelse(all(is.na(x)), NA, sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x))))
)

sg_summary_plot <- seagrass_clean %>% 
  group_by(Date,Plot_Label,Plot_Type) %>% 
  summarize(across(c(Percent_Cover,Canopy_Height_cm),
                   summary_calcs))


sg_summary_plot_type <- sg_summary_plot %>% 
  group_by(Date,Plot_Type) %>% 
  summarize(across(c(Percent_Cover_Mean,Canopy_Height_cm_Mean),
                   summary_calcs))
  
```




#### Oyster Data - Abundance
```{r}
#| echo: false

# 
# summary_calcs <- list(
#   Mean = function(x) mean(x, na.rm = TRUE),
#   n = function(x) sum(!is.na(x)),
#   SD = function(x) ifelse(all(is.na(x)), NA, sd(x, na.rm = TRUE)),
#   SE = function(x) ifelse(all(is.na(x)), NA, sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x))))
# )

oyst_abundance_summary_tile <- oyster_abundance_clean %>% 
  group_by(Date,
           Plot_Label,
           Plot_Type,
           Tile_Type,
           Tile_Tag) %>% 
  summarize(across(c(Oyster_Abundance),
                   summary_calcs)) %>% 
  ungroup()

# summarize the oyster data by plot type but SEPARATE for the predator versus no-predator cue tiles
oyst_abundance_summary_plot <- oyst_abundance_summary_tile %>% 
  group_by(Date,
           Plot_Label,
           Plot_Type,
           Tile_Type) %>% 
  summarize(across(c(Oyster_Abundance_Mean),
                   summary_calcs)) %>% 
  group_by(Date,
           Plot_Type,
           Tile_Type) %>% 
  summarize(across(c(Oyster_Abundance_Mean_Mean),
                   summary_calcs))

# summarize the oyster data by plot type but for all 4 tiles, regardless of predator treatment (predator vs. no-predator cues)
oyst_abundance_summary_plot_type <- oyst_abundance_summary_tile %>% 
  group_by(Date,
           Plot_Label,
           Plot_Type) %>% 
  summarize(across(c(Oyster_Abundance_Mean),
                   summary_calcs)) %>% 
  group_by(Date,
           Plot_Type) %>% 
  summarize(across(c(Oyster_Abundance_Mean_Mean),
                   summary_calcs))

```




#### Oyster Data - Size
```{r}
#| echo: false

# 
# summary_calcs <- list(
#   Mean = function(x) mean(x, na.rm = TRUE),
#   n = function(x) sum(!is.na(x)),
#   SD = function(x) ifelse(all(is.na(x)), NA, sd(x, na.rm = TRUE)),
#   SE = function(x) ifelse(all(is.na(x)), NA, sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x))))
# )

oyst_size_summary_tile <- oyster_size_clean %>% 
  group_by(Date,
           Plot_Label,
           Plot_Type,
           Tile_Type,
           Tile_Tag) %>% 
  summarize(across(c(Oyster_Length_mm),
                   summary_calcs)) %>% 
  ungroup()

# summarize the oyster data by plot type but SEPARATE for the predator versus no-predator cue tiles
oyst_size_summary_plot <- oyst_size_summary_tile %>% 
  group_by(Date,
           Plot_Label,
           Plot_Type,
           Tile_Type) %>% 
  summarize(across(c(Oyster_Length_mm_Mean),
                   summary_calcs)) %>% 
  group_by(Date,
           Plot_Type,
           Tile_Type) %>% 
  summarize(across(c(Oyster_Length_mm_Mean_Mean),
                   summary_calcs))

# summarize the oyster data by plot type but for all 4 tiles, regardless of predator treatment (predator vs. no-predator cues)
oyst_size_summary_plot_type <- oyst_size_summary_tile %>% 
  group_by(Date,
           Plot_Label,
           Plot_Type) %>% 
  summarize(across(c(Oyster_Length_mm_Mean),
                   summary_calcs)) %>% 
  group_by(Date,
           Plot_Type) %>% 
  summarize(across(c(Oyster_Length_mm_Mean_Mean),
                   summary_calcs)) %>% 
  ungroup()

```




### Plotting data in figures

#### Seagrass Data Figures
```{r}
#| echo: false
#| eval: false

ggplot(data = sg_summary_plot,
       aes(x = Plot_Type, y = Percent_Cover_Mean)) +
  facet_wrap(~Date, nrow = 1, ncol = 2,
             labeller = labeller(Date = c("7/31/2023" = "Initial",
                                          "9/6/23" = "One Month Post-Deployment")),
             scales = "free") +
  theme_classic() + 
  geom_errorbar(aes(ymin = Percent_Cover_Mean - Percent_Cover_SE,
                     ymax = Percent_Cover_Mean + Percent_Cover_SE),
                width = 0.2,
                position = position_jitter(width = 0.3, seed = 123)) +
  geom_point(position = position_jitter(width = 0.3, seed = 123),
             size = 4,
             aes(shape = "plot average", fill = "plot average", color = "plot average"),
             show.legend = TRUE) +
  geom_errorbar(data = sg_summary_plot_type,
                inherit.aes = FALSE,
                aes(x = Plot_Type,
                    ymin = Percent_Cover_Mean_Mean - Percent_Cover_Mean_SE,
                    ymax = Percent_Cover_Mean_Mean + Percent_Cover_Mean_SE),
                 width = 0.2) +
  geom_point(data = sg_summary_plot_type,
             inherit.aes = FALSE,
             aes(x = Plot_Type, y = Percent_Cover_Mean_Mean,
                 shape = "plot type average", fill = "plot type average", color = "plot type average"),
             size = 6,
             show.legend = TRUE) +
  ylim(0,100) +
  labs(x = "Plot Type",
       y = "Seagrass Percent Cover (mean±SE)") +
  scale_color_manual(values = c("plot average" = "black", "plot type average" = "black"), name = "Legend") +
  scale_shape_manual(values = c("plot average" = 21, "plot type average" = 23), name = "Legend") +
  scale_fill_manual(values = c("plot average" = "lightgreen", "plot type average" = "darkgreen"), name = "Legend") +
  guides(
    shape = guide_legend(override.aes = list(size = 4))
  ) +
  theme(panel.border = element_blank())

```



```{r}
#| echo: false


## function to easily plot the seagrass percent cover and canopy height data without having to copy and paste and rename all the variables....

create_custom_ggplot <- function(summary1, summary2, variable_name, y_axis_label, y_axis_min, y_axis_max) {
  gg <- ggplot(summary1, aes(x = Plot_Type, y = !!sym(paste0(variable_name, "_Mean")))) +
    facet_wrap(~Date, nrow = 1, ncol = 2,
               labeller = labeller(Date = c("7/31/2023" = "Initial", "9/6/23" = "One Month Post-Deployment")),
               scales = "free") +
    theme_classic() + 
    geom_errorbar(aes(ymin = !!sym(paste0(variable_name, "_Mean")) - !!sym(paste0(variable_name, "_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean")) + !!sym(paste0(variable_name, "_SE"))),
                  width = 0.2,
                  position = position_jitter(width = 0.3, seed = 123)) +
    geom_point(position = position_jitter(width = 0.3, seed = 123),
               size = 4,
               aes(shape = "plot average", fill = "plot average", color = "plot average"),
               show.legend = TRUE) +
    geom_errorbar(data = summary2,
                  inherit.aes = FALSE,
                  aes(x = Plot_Type,
                      ymin = !!sym(paste0(variable_name, "_Mean_Mean")) - !!sym(paste0(variable_name, "_Mean_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean_Mean")) + !!sym(paste0(variable_name, "_Mean_SE"))),
                  width = 0.2) +
    geom_point(data = summary2,
               inherit.aes = FALSE,
               aes(x = Plot_Type, y = !!sym(paste0(variable_name, "_Mean_Mean")),
                   shape = "plot type average", fill = "plot type average", color = "plot type average"),
               size = 6,
               show.legend = TRUE) +
    ylim(y_axis_min, y_axis_max) +
    labs(x = "Plot Type",
         y = paste(y_axis_label, " (mean±SE)")) +
    scale_color_manual(values = c("plot average" = "black", "plot type average" = "black"), name = "Legend") +
    scale_shape_manual(values = c("plot average" = 21, "plot type average" = 23), name = "Legend") +
    scale_fill_manual(values = c("plot average" = "lightgreen", "plot type average" = "darkgreen"), name = "Legend") +
    guides(shape = guide_legend(override.aes = list(size = 4)))+
    theme(panel.border = element_blank())
  
  return(gg)
}



seagrass_percent_cover_fig <- create_custom_ggplot(sg_summary_plot, sg_summary_plot_type, "Percent_Cover", "Seagrass Percent Cover", 0, 100)
print(seagrass_percent_cover_fig)

seagrass_canopy_height_fig <- create_custom_ggplot(sg_summary_plot, sg_summary_plot_type, "Canopy_Height_cm", "Seagrass Canopy Height (cm)", 0, 25)
print(seagrass_canopy_height_fig)


```


#### Oyster Data Figures

```{r providing an x-axis jitter value for all oyster datasets}

# jitter_values <- function(){
#   
# }
  
oyst_abundance_summary_tile <- oyst_abundance_summary_tile %>% 
  mutate(x_jitter = ifelse(Plot_Type == "Seagrass+Tiles", 0.18, 0.68)) %>% 
  mutate(x_jitter = ifelse(Tile_Type == "Predator", x_jitter + 0.14, x_jitter)) %>% 
  mutate(Tile_Type = as.factor(Tile_Type))

oyst_abundance_summary_plot <- oyst_abundance_summary_plot %>% 
  mutate(x_jitter = ifelse(Plot_Type == "Seagrass+Tiles", 0.21, 0.71)) %>% 
  mutate(x_jitter = ifelse(Tile_Type == "Predator", x_jitter + 0.14, x_jitter)) %>% 
  mutate(Tile_Type = as.factor(Tile_Type))

oyst_abundance_summary_plot_type <- oyst_abundance_summary_plot_type %>% 
  mutate(x_jitter = ifelse(Plot_Type == "Seagrass+Tiles", 0.25, 0.75))

```



```{r custom function for plotting oyster data}


oyster_custom_ggplot <- function(summary1, summary2, summary3, variable_name, y_axis_label, y_axis_min, y_axis_max) {
  gg <- ggplot(summary1, aes(x = x_jitter, y = !!sym(paste0(variable_name, "_Mean")))) +
    facet_wrap(~Date, nrow = 1, ncol = 2,
               labeller = labeller(Date = c("8/2/23" = "Initial", "9/6/23" = "One Month Post-Deployment")),
               scales = "free") +
    theme_classic() + 
    geom_errorbar(aes(ymin = !!sym(paste0(variable_name, "_Mean")) - !!sym(paste0(variable_name, "_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean")) + !!sym(paste0(variable_name, "_SE"))),
                  width = 0.1) +
    geom_point(size = 4,
               aes(shape = "tile average", fill = "tile average", color = "tile average"),
               show.legend = TRUE) +
    geom_errorbar(data = summary2,
                  inherit.aes = FALSE,
                  aes(x = x_jitter,
                      ymin = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) - !!sym(paste0(variable_name, "_Mean_Mean_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) + !!sym(paste0(variable_name, "_Mean_Mean_SE"))),
                  width = 0.1) +
    geom_point(data = summary2,
               inherit.aes = FALSE,
               aes(x = x_jitter, y = !!sym(paste0(variable_name, "_Mean_Mean_Mean")),
                   shape = "plot average", fill = "plot average", color = "plot average"),
               size = 6,
               show.legend = TRUE) +
    geom_errorbar(data = summary3,
                  inherit.aes = FALSE,
                  aes(x = x_jitter,
                      ymin = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) - !!sym(paste0(variable_name, "_Mean_Mean_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) + !!sym(paste0(variable_name, "_Mean_Mean_SE"))),
                  width = 0.1) +
    geom_point(data = summary3,
               inherit.aes = FALSE,
               aes(x = x_jitter, y = !!sym(paste0(variable_name, "_Mean_Mean_Mean")),
                   shape = "plot type average", fill = "plot type average", color = "plot type average"),
               size = 6,
               show.legend = TRUE) +
    scale_x_continuous(breaks = c(0.25, 0.75), labels = c("Seagrass+Tiles", "Tiles")) +
    ylim(y_axis_min, y_axis_max) +
    labs(x = "Plot Type",
         y = paste(y_axis_label, " (mean±SE)")) +
    scale_color_manual(values = c("tile average" = "black", "plot average" = "black", "plot type average" = "black"), name = "Legend") +
    scale_shape_manual(values = c("tile average" = 21, "plot average" = 21, "plot type average" = 23), name = "Legend") +
    scale_fill_manual(values = c("tile average" = "lightblue", "plot average" = "darkblue", "plot type average" = "gold"), name = "Legend") +
    guides(shape = guide_legend(override.aes = list(size = 4)))+
    theme(panel.border = element_blank())
  
  return(gg)
}


oyster_custom_ggplot(oyst_abundance_summary_tile,
                     oyst_abundance_summary_plot,
                     oyst_abundance_summary_plot_type,
                     "Oyster_Abundance",
                     "Oyster Abundance",
                     0, 35)

```



```{r}
## revised oyster ggplot function from chatGPT suggestions

oyster_custom_ggplot_2 <- function(summary1, summary2, summary3, variable_name, y_axis_label, y_axis_min, y_axis_max) {
  gg <- ggplot(summary1, aes(x = x_jitter, y = !!sym(paste0(variable_name, "_Mean")))) +
    facet_wrap(~Date, nrow = 1, ncol = 2,
               labeller = labeller(Date = c("8/2/23" = "Initial", "9/6/23" = "One Month Post-Deployment")),
               scales = "free") +
    theme_classic() + 
    geom_errorbar(aes(ymin = !!sym(paste0(variable_name, "_Mean")) - !!sym(paste0(variable_name, "_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean")) + !!sym(paste0(variable_name, "_SE"))),
                  width = 0.1) +
    geom_point(size = 4, aes(shape = "tile average", fill = Tile_Type, color = Tile_Type),
               show.legend = TRUE) +
    scale_color_manual(values = c("Level1" = "red", "Level2" = "blue", "Level3" = "green")) +
    scale_fill_manual(values = c("Level1" = "red", "Level2" = "blue", "Level3" = "green")) +
    # 
    # geom_point(size = 4,
    #            aes(shape = "tile average", fill = Tile_Type, color = Tile_Type),
    #            show.legend = TRUE) +
    geom_errorbar(data = summary2,
                  inherit.aes = FALSE,
                  aes(x = x_jitter,
                      ymin = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) - !!sym(paste0(variable_name, "_Mean_Mean_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) + !!sym(paste0(variable_name, "_Mean_Mean_SE"))),
                  width = 0.1) +
    geom_point(data = summary2,
               inherit.aes = FALSE,
               aes(x = x_jitter, y = !!sym(paste0(variable_name, "_Mean_Mean_Mean")),
                   shape = "plot average", fill = Tile_Type, color = Tile_Type),
               size = 6,
               show.legend = TRUE) +
    geom_errorbar(data = summary3,
                  inherit.aes = FALSE,
                  aes(x = x_jitter,
                      ymin = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) - !!sym(paste0(variable_name, "_Mean_Mean_SE")),
                      ymax = !!sym(paste0(variable_name, "_Mean_Mean_Mean")) + !!sym(paste0(variable_name, "_Mean_Mean_SE"))),
                  width = 0.1) +
    geom_point(data = summary3,
               inherit.aes = FALSE,
               aes(x = x_jitter, y = !!sym(paste0(variable_name, "_Mean_Mean_Mean")),
                   shape = "plot type average", fill = "plot type average", color = "plot type average"),
               size = 6,
               show.legend = TRUE) +
    scale_x_continuous(breaks = c(0.25, 0.75), labels = c("Seagrass+Tiles", "Tiles")) +
    ylim(y_axis_min, y_axis_max) +
    labs(x = "Plot Type",
         y = paste(y_axis_label, " (mean±SE)")) +
    scale_color_manual(values = c("tile average" = "black", "plot average" = "black", "plot type average" = "black",
                                  Tile_Type = c("Level1" = "gray", "Level2" = "blue", "Level3" = "yellow")), name = "Legend") +
    scale_shape_manual(values = c("tile average" = 21, "plot average" = 21, "plot type average" = 23), name = "Legend") +
    scale_fill_manual(values = c("tile average" = "lightblue", "plot average" = "darkblue", "plot type average" = "gold"), name = "Legend") +
    # 
    # scale_color_manual(values = c("tile average" = "black", "plot average" = "black", "plot type average" = "black"), name = "Legend") +
    # scale_shape_manual(values = c("tile average" = 21, "plot average" = 21, "plot type average" = 23), name = "Legend") +
    # scale_fill_manual(values = c("tile average" = "lightblue", "plot average" = "darkblue", "plot type average" = "gold"), name = "Legend") +
    guides(shape = guide_legend(override.aes = list(size = 4)))+
    theme(panel.border = element_blank())
  
  return(gg)
}


oyster_custom_ggplot_2(oyst_abundance_summary_tile,
                     oyst_abundance_summary_plot,
                     oyst_abundance_summary_plot_type,
                     "Oyster_Abundance",
                     "Oyster Abundance",
                     0, 35)

```



