---
title: "GROW DISL Report"
author: "Shayna A. Sura"
format: pdf
editor: visual
---

<!-- To create this document, must first run the associated .qmd file named "grow-2023-data.qmd". -->

# GROW Project

# Dauphin Island Sea Lab

# Phase 2

## Update Report

## Data Summary

### Oysters

```{r}
#| echo: false
#| warning: false
#| message: false


```

**Figure 1.** The abundance of oysters on Tiles a) before deployment (initials) and b) one-month post-deployment into the field. Tiles were deployed into either Seagrass+Tiles plots or Tiles Only plots.

```{r}
#| echo: false
#| warning: false
#| message: false


```

**Figure 2.** The size of oysters (mm) on Tiles a) before deployment (initials) and b) one-month post-deployment into the field. Tiles were deployed into either Seagrass+Tiles plots or Tiles Only plots.

### Seagrass

```{r}
#| echo: false
#| warning: false
#| message: false

# {{< include _grow-2023-data.qmd >}}
# 
# {{< include _grow-2023-data.qmd>}}

```

![](images/Fig_3_seagrass_percent_cover.pdf){fig-align="center" width="4in" height="6in"}

**Figure 3.** The percent cover of seagrass in Seagrass and Seagrass+Tiles plot types a) before deployment of Tiles (initials) and b) one-month post-deployment of Tiles into the field. Seagrass percent cover was assessed in 4 quadrats per plot. There are three replicate plots per plot type (Seagrass, Seagrass+Tiles).

```{r}
#| echo: false
#| warning: false
#| message: false


# Fig_2_Seagrass_Canopy_Height
```

![](images/Fig_4_seagrass_canopy_height.pdf){fig-align="center" width="4in" height="6in"}

**Figure 4.** The canopy height of seagrass (cm) in Seagrass and Seagrass+Tiles plot types a) before deployment of Tiles (initials) and b) one-month post-deployment of Tiles into the field. Seagrass percent cover was assessed in 4 quadrats per plot. There are three replicate plots per plot type (Seagrass, Seagrass+Tiles). NOTE: seagrass canopy height was not measured at the one-month monitoring time point.
